---
title: "Permutation Tests and Variable Selection"
author: "Nina Zumel"
date: "March 16, 2016"
output:
  md_document:
    variant: markdown_github
---



For simplicity, we will do this all with glm, and mostly with deviance,
but in theory this method will determine model significance for any
modeling method and any metric. However, for variable selection,
you might as well use a simple and easy to run modeling method like
glm or lm, regardless of what your final modeling algorithm will be.
In fact, permutation tests with lm can be made very efficient, because
you are keeping the same design matrix and only changing the y vector.

```{r modelSignificance}
source("functions.R")
# install.packages("devtools")
# devtools::install_github("WinVector/WVPlots",build_vignettes=TRUE)
library("WVPlots")

set.seed(12959437)
```

## Some Examples of Estimating Model Significance

### A case with signal
```{r cleandata}
# clean data
run_example(ngood=10, nnoise=3,
            datasize=1000, nperm=500, 'Data with signal')
```

### A case with no signal
```{r nosignal}
run_example(ngood=0, nnoise=10,
            datasize=1000, nperm=500, 'Data with no signal')
```

### Bad Bayes: very wide data with no signal
```{r badbayes}
run_example(ngood=0, nnoise=300, 
            datasize=1000, nperm=200, 'Bad Bayes situation')
```

## Using model significance (chi-squared test) to select variables.

### Data with signal
```{r varSelclean}
run_vs_example(ngood=10, nnoise=20,
               datasize=1000, nperm=200,
               threshold=0.05, 'Data with signal, threshold p=0.05')
```

### Wider data with some signal
```{r varSelnoisySig}
run_vs_example(ngood=5, nnoise=100,
               datasize=1000, nperm=200,
               threshold=0.05, 'Wide data with some signal, threshold p=0.05')
```

### Data with no signal

You would expect a threshold of 0.05 (or 1/20) to let about one out of every twenty noise variables to slip by. This gives you a rule of thumb for picking the threshold.

```{r varSelnosig}
# scoring columns, data with no signal
nnoise = 100
p=0.05
# the number of bad variables to expect
print(nnoise*p)

run_vs_example(ngood=0, nnoise=nnoise,
               datasize=1000, nperm=200,
               threshold=p, 'Data with no signal, threshold p=0.05')


```

### Bad Bayes
```{r varSelnosig2}
nnoise=300
p=0.01
# the number of bad variables to expect
print(nnoise*p)

run_vs_example(ngood=0, nnoise=nnoise, 
            datasize=1000, nperm=200, 
            threshold=p, 'Bad Bayes situation, threshold p=0.01')

```
